### 2021-05-24 assertions/doc/thesis-bernhardt/chapters/6-conclusion.tex
Final revision - conclusion + appendix

\chapter{Conclusion and Outlook}\label{chap:conclusion}
Complex blockchain-based applications suffer from high computation costs and limited scalability. A consequential measure to address this issue, is to move computations off the chain and only verify the correctness of submitted results on the blockchain. This thesis  proposed assertions for distributed refutation as a probabilistic implementation of this verification. In the processing model of this approach, the validators of the system are challenged to assert specified properties for a random subset of the off-chain computation. If an assertion fails, the respective validator generates a counterexample and publishes it to the network. When the system acknowledges the counterexample, the off-chain computation is rejected. Suitable use-cases for this approach are computations, whose properties can be specified with formulas in predicate logic featuring universal quantification. To support this model on the target platform Tezos, a toolchain was developed to allow developers to specify explicit assertions and generate corresponding smart contracts. These contracts check the property point-wise within the quantification domain and identify counterexamples when executed by the validators. For the specification of assertions, a simple domain-specific language was defined.

For the proposed architecture, the conducted costs analysis shows high transaction fees for the submission of off-chain computations,  due to high verification costs. Although the verification might still be more efficient than the execution itself, it can be assumed that the cost reductions are equalised in cases of moderate complexity of off-chain computations. Inter-contract communication was identified as a main driving factor of the verification cost. To lower these costs and approach usability on production systems, the thesis provided possible improvements and modifications of the architecture.

@@

Based on either the proposed or a modification of the architecture, subsequent tasks include the implementation of the compiler in order to conclude the toolchain. As addressed in the introduction, the implementation of distributed assertions on blockchains also requires a design and implementation of a protocol amendment. Many of the aspects that need to be considered by the protocol design have already been pointed out by this work. This includes, i.a., the extensions to Michelson and Tezos' VM to support random value generation, the implementation of a dedicated execution mode, the withholding of the transaction submitting an off-chain computation until it has been verified, and the rejection of this transaction as a consequence of a verified counterexample.

Besides the completion of the proof-of-concept implementation, the assertion language can be extended with features, such as variables and user-defined functions, to improve (re-)usability. Furthermore, an interesting extension is the support for properties formulated with existential quantification, as this allows to verify off-chain computations by distributed validation in addition to distributed refutation.

>>>>

\chapter{Conclusion and Outlook}\label{chap:conclusion}
Complex blockchain-based applications suffer from high computation costs and limited scalability. A consequential measure to address this issue is to move computations off the chain and only verify the correctness of submitted results on the blockchain. This thesis  proposed assertions for distributed refutation as a probabilistic implementation of this verification. In the processing model of this approach, the validators of the system are challenged to assert specified properties for a random subset of the off-chain computation. If an assertion fails, the respective validator generates a counterexample and publishes it to the network. When the system acknowledges the counterexample, the off-chain computation is rejected. Suitable use-cases for this approach are computations, whose properties can be specified with formulas in predicate logic featuring universal quantification. To support this model on the target platform Tezos, a toolchain was developed to allow developers to specify explicit assertions and generate corresponding smart contracts. These contracts check the property point-wise within the quantification domain and can identify counterexamples with a configurable probability when executed by the validators. For the specification of assertions, a simple domain-specific language was defined.

Based on the proposed architecture, the conducted costs analysis shows high transaction fees for the submission of off-chain computations due to high verification costs. Although the verification might still be more efficient than the execution itself, it can be assumed that the cost reductions are not significant enough in cases of low or moderate complexity of off-chain computations. Inter-contract communication was identified as a main driving factor of the verification cost. To lower these costs and approach usability on production systems, the thesis provided possible improvements and modifications of the architecture.

@@

Based on the preparations and work done in this thesis, subsequent tasks on this project include the implementation of the compiler to conclude the toolchain. As addressed in the introduction, the implementation of distributed assertions on blockchains also requires a design and implementation of a protocol amendment. Many of the aspects that need to be considered by the protocol design have already been pointed out throughout this thesis. This includes, i.a., the extensions to Michelson and Tezos' VM to support random value generation, the implementation of a dedicated execution mode, the withholding of the transaction submitting an off-chain computation until the assertions have been executed, and the rejection of this transaction as a consequence of a verified counterexample.

Besides the completion of the proof-of-concept implementation, the assertion language can be extended with features to improve its (re-)usability, such as variables and user-defined functions. Furthermore, an interesting extension to the execution model is the support for properties formulated with existential quantification, as this allows to verify off-chain computations by distributed validation as opposed to to distributed refutation.

### 2021-05-24 assertions/doc/thesis-bernhardt/chapters/appendix.tex
Final revision - conclusion + appendix

\appendix
\chapter{Michelson Implementations of \texttt{Foo} and \texttt{Bar}}\label{apx:cost_analysis_contract}
The following contracts are the Michelson equivalent of the dummy contracts given in \secref{sec:usecase_cost}.
\section{Foo}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 caption=Dummy Michelson contract executing a dynamic loop]{listings/foo.tz}

\section{Bar}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 caption=Dummy Michelson contract iterating over a dynamic list]{listings/bar.tz}

\chapter{Assertion Grammar in EBNF}\label{apx:grammar}
The following two versions of the grammar implement the prefix and infix notations for the assertion syntax.

@@

@@

\chapter{Controller Contract Code in Liquidity}\label{apx:controller_liq}
\lstinputlisting[label=lst:manager_templ_liq, language=Michelson, caption=Controller code in Liquidity]{listings/controller.liq}

\chapter{Controller Contract Code in Michelson}\label{apx:controller_michelson}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:controller_michelson,
				 caption=Controller code in Michelson]{listings/controller.tz}

\chapter{Distributed Michelson Implementation of \texttt{Foo} and \texttt{Bar}}\label{apx:foobar_distributed}
The following contracts are a restructuring of the contracts given in Appendix \ref{apx:cost_analysis_contract}, s.t. they implement the proposed orchestration strategy.

\section{Foo}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:foo_manager,
				 caption=Controller contract of \texttt{Foo}]{listings/foo_manager.tz}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:foo_assertion,
				 caption=Assertion contract of \texttt{Foo}]{listings/foo_assertion.tz}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:foo_parent,
				 caption=Parent contract of \texttt{Foo}]{listings/foo_parent.tz}

\section{Bar}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:bar_manager,
				 caption=Controller contract of \texttt{Bar}]{listings/bar_manager.tz}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:bar_assertion,
				 caption=Assertion contract of \texttt{Bar}]{listings/bar_assertion.tz}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:bar_parent,
				 caption=Parent contract of \texttt{Bar}]{listings/bar_parent.tz}

>>>>

\appendix
\chapter{Michelson Implementations of \texttt{Foo} and \texttt{Bar}}\label{apx:cost_analysis_contract}
The following contracts are the Michelson implementations of the dummy contracts given in \secref{sec:usecase_cost}.
\section{Foo}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 caption=Dummy Michelson contract executing a dynamic loop (\texttt{Foo.tz})]{listings/foo.tz}

\section{Bar}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 caption=Dummy Michelson contract iterating over a dynamic list (\texttt{Bar.tz})]{listings/bar.tz}

\chapter{Assertion Grammar in EBNF}\label{apx:grammar}
The following two versions of a grammar in extended Backus-Naur form (EBNF) implement the prefix and infix notations for the assertion syntax.

@@

The following listings show the implementation of the list accessing operator \texttt{nth} using the pre-defined loop instruction. \lstref{lst:nth_liq} shows a readable implementation in Liquidity; it contains necessary boilerplate code to make it compilable. The compilation result (including the boilerplate code) is given in \lstref{lst:nth_liq}.

@@

\chapter{Controller Contract Code in \\ \mbox{Liquidity}}\label{apx:controller_liq}
\lstinputlisting[label=lst:manager_templ_liq, language=Michelson, caption=Controller code in Liquidity]{listings/controller.liq}

\chapter{Controller Contract Code in Michelson}\label{apx:controller_michelson}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:controller_michelson,
				 caption=Controller code in Michelson featuring one gateway and one assertion entrypoint]{listings/controller.tz}

\chapter{Distributed Michelson Implementation of \texttt{Foo} and \texttt{Bar}}\label{apx:foobar_distributed}
The following contracts are a restructuring of the contracts given in Appendix \ref{apx:cost_analysis_contract}, s.t. they implement the modular orchestration scheme.

\section{Contract Foo}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:foo_manager,
				 caption=Controller contract for \texttt{Foo.tz}]{listings/foo_manager.tz}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:foo_assertion,
				 caption=Assertion contract for \texttt{Foo.tz}]{listings/foo_assertion.tz}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:foo_parent,
				 caption=Parent contract (\texttt{Foo.tz})]{listings/foo_parent.tz}

\section{Contract Bar}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:bar_manager,
				 caption=Controller contract for \texttt{Bar.tz}]{listings/bar_manager.tz}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:bar_assertion,
				 caption=Assertion contract for \texttt{Bar.tz}]{listings/bar_assertion.tz}
\lstinputlisting[basicstyle=\linespread{1.0}\fontfamily{lmr}\selectfont\small,
				 backgroundcolor=\color{cverbbg},
				 linewidth=14cm,
				 xleftmargin=0.5cm,
				 frame=lr,
				 framesep=8pt,
				 framerule=0pt,
				 captionpos=b,
				 numbers=none,
				 language=Michelson,
				 label=lst:bar_parent,
				 caption=Parent contract (\texttt{Bar.tz})]{listings/bar_parent.tz}
